---
title: "HW 2B, Principles of Statistical Graphs - Nikhil Kotecha"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Goal is to get as close to this economist graph as possible (maybe even improve on it)! Here's a link with the orignial: http://www.economist.com/node/21541178

I relied on a harvard tutorial to help me get up to speed with ggplot2:http://tutorials.iq.harvard.edu/R/Rgraphics/Rgraphics.html#orgbe01a81

(I have working knowledge of R from working with data for neural nets but mainly I use python). 

Data comes from the linked tutorial with information on country name, Human Development Index, Corruption Perceptions Index, and associated region.  

I slowly added information, and found that the most interesting use of space was to show HDI by CPI clustered by region. I initially tried to show all the regions in one graph, but it was too messy. Instead I graphed three regions separately (to stay in the 4 graph limit.) The hypotehtical audience for the plot's are a lay audience interested in new from around the world. My goals in making the graph were to learn how to replicate an economist-esque graph. I was particularly interested in trying a graphic in the vein of a mass media outlet after going through the "Visual Revelations: Improving Data" article in this week's reading. I have tried academic graphics before, but never through R and wanted to bolster my skills here too. In each graph the comparisions are similar: I wanted to show regional differences in perceived corruption and the development of the nation. Initially it was to compare across all regions (the first graph), but later graphs I think more effective because they compare within region. 

Code Below:

```{r}
library(ggplot2)
library(ggrepel)

EconomistData <- read.csv(file="~/Desktop/EconomistData.csv",head=TRUE,sep=",")

p1 <- ggplot(EconomistData, aes(x = CPI, y = HDI, color=Region)) + geom_point()

p2 <- p1 + geom_smooth(method = loess, fill =NA)

p3 <- p2 + labs(x = "Corruption Perceptions Index, 2011 (10=least corrupt)", y ="Human Development Index, 2011 (1=best)", title ="Corruption and Human Development", subtitle = "All Regions")

p4 <- p3 + geom_text(aes(label=Country), size = 3)

#p5 <-p4 + facet_wrap(~Region, ncol = 10)
#p5 was super hard to read, so broke into separate graphs

Asia_Pacific <-
  ggplot(subset(EconomistData, Region %in% c("Asia Pacific")),
  aes(x = CPI,
  y = HDI,
  label=Region)) +
  geom_point(color="brown") +
  geom_smooth(method = loess, fill = NA, color="brown") +
  labs(x = "Corruption Perceptions Index, 2011 (10=least corrupt)", y =
  "Human Development Index, 2011 (1=best)", title = "Corruption and Human Development", subtitle = "Asia Pacific")+
  geom_text_repel(aes(label=Country), size = 3)

Americas <- ggplot(subset(EconomistData, Region %in% c("Americas")),
       aes(x=CPI,
           y=HDI,
           label=Region))+
  geom_point(color="blue")+
  geom_smooth(method = loess, fill =NA, color="blue")+
  labs(x = "Corruption Perceptions Index, 2011 (10=least corrupt)", y ="Human Development Index, 2011 (1=best)", title ="Corruption and Human Development", subtitle = "Americas")+ 
  geom_text_repel(aes(label=Country), size = 3)
  
SSA <- ggplot(subset(EconomistData, Region %in% c("SSA")),
       aes(x=CPI,
           y=HDI,
           label=Region))+
  geom_point(color="pink")+
  geom_smooth(method = loess, fill =NA, color="pink")+
  labs(x = "Corruption Perceptions Index, 2011 (10=least corrupt)", y ="Human Development Index, 2011 (1=best)", title ="Corruption and Human Development", subtitle = "Sub-Saharan Africa")+ 
  geom_text_repel(aes(label=Country), size = 3)

```

## Plots

```{r pressure, echo=FALSE}

p4

Asia_Pacific

Americas

SSA

```