---
title: "Hw 8a/8b"
authors: "Tin Oreskovic and Nikhil Kotecha"
date: "Static: 10/24/2017, Dynamic: 10/29/2017 "
output:
  html_document: default
  pdf_document: default
---
Sources: http://kateto.net/network-visualization

```{r setup, include=FALSE}

library(readr)
library(igraph)
library(network)
library(visNetwork)
library(sna)
library(threejs)
library(networkD3)
library(ndtv)

nodes1 <- read_csv("~/Desktop/Uber Folder/Columbia/Data sets/Network Visualization - polnet2017/Data files/Dataset1-Media-Example-NODES.csv")
links1 <- read_csv("~/Desktop/Uber Folder/Columbia/Data sets/Network Visualization - polnet2017/Data files/Dataset1-Media-Example-EDGES.csv")

#nodes1 <- Dataset1_Media_Example_NODES
#links1 <- Dataset1_Media_Example_EDGES


nodes2 <- read_csv("~/Desktop/Uber Folder/Columbia/Data sets/Network Visualization - polnet2017/Data files/Dataset2-Media-User-Example-NODES.csv")
links2 <- read_csv("~/Desktop/Uber Folder/Columbia/Data sets/Network Visualization - polnet2017/Data files/Dataset2-Media-User-Example-EDGES.csv") 

#nodes2 <- Dataset2_Media_User_Example_NODES
#links2 <- Dataset2_Media_User_Example_EDGES

links2 <- as.matrix(links2)
dim(links2)
dim(nodes2)

net2 <- graph_from_incidence_matrix(links2)
table(V(net2)$type)

net <- graph_from_data_frame(d=links1, vertices=nodes1, directed=T)
# Get an edge list or a matrix:
as_edgelist(net, names=T)
as_adjacency_matrix(net, attr="weight")

# Or data frames describing nodes and edges:
as_data_frame(net, what="edges")
as_data_frame(net, what="vertices")

net2 <- graph_from_incidence_matrix(links2)
table(V(net2)$type)

```

## Different Approaches with fake data

A couple of ways to potentially visualize data: cirlce, and with the Fruchterman-Reingold Algorithm. 

I like the FR algo because it offers a nice intuitive understanding of the data. 


```{r cars}
#Fake data, trying out different algorithms
net.bg <- sample_pa(80) 
V(net.bg)$size <- 8
V(net.bg)$frame.color <- "white"
V(net.bg)$color <- "orange"
V(net.bg)$label <- "" 
E(net.bg)$arrow.mode <- 0
plot(net.bg)

#circle - kinda hard to see
l <- layout_in_circle(net.bg)
plot(net.bg, layout=l)


#Fruchterman-Reingold: force-directed layout algo
## edges are similar, and cross minimally
### graph as physical system, with electrical repulsion/acting as springs
#non-deterministic
fr <- layout_with_fr(net.bg)
plot(net.bg, layout=fr)

```

## Plot with Newspaper data

Trying the Newspaper data without any fancy algorithm to sort the information.  I just wanted to get a feel for the information. In the dynamic graph, I will try to show how the network evolves over time. 

```{r pressure, echo=TRUE}

#actual data to play around 
# Plot with curved edges (edge.curved=.1) and reduce arrow size:
# Note that using curved edges will allow you to see multiple links
# between two nodes (e.g. links going in either direction, or multiplex links)

net <- simplify(net, remove.multiple = F, remove.loops = T) 
plot(net, edge.arrow.size=.4,vertex.label=NA)
plot(net, edge.arrow.size=.4, edge.curved=.1)

# Set edge color to light gray, the node & border color to orange 
# Replace the vertex label with the node names stored in "media"
plot(net, edge.arrow.size=.2, edge.color="orange",
     vertex.color="orange", vertex.frame.color="#ffffff",
     vertex.label=V(net)$media, vertex.label.color="black") 


```

## Dynamic graph with networkD3

I've used shiny before, so wanted to see what D3 could do for network visualization. (Comments in code.)

```{r}

library(networkD3)

#make nodes numeric and starting from 0. 
el <- data.frame(from=as.numeric(factor(links1$from))-1, 
                 to=as.numeric(factor(links1$to))-1 )

#nodes need to be in same order as the source column in links. 
nl <- cbind(idn=factor(nodes1$media, levels=nodes1$media), nodes1)

# interactive chart: group = color; nodesize = (confusingly) number of column in node data, used for sizing; (based on a similar algorithm to above) charge = if negative, refers to node repulsion OR if positive, attraction. 

#et voila
forceNetwork(Links = el, Nodes = nl, Source="from", Target="to",
               NodeID = "idn", Group = "type.label",linkWidth = 1,
               linkColour = "#afafaf", fontSize=12, zoom=T, legend=T,
               Nodesize=6, opacity = 0.8, charge=-300, 
               width = 600, height = 400)
```

## Dynamic Graph VisNetwork

This is kinda fun, so let's check out some other r packages and see what they can do. 

```{r}

library('visNetwork') 

#basic visualization
visNetwork(nodes1, links1, width="100%", height="400px", main="Network")

nodes1$shape <- "dot"  
nodes1$shadow <- TRUE # Nodes will drop shadow
nodes1$title <- nodes1$media # Text on click
nodes1$label <- nodes1$type.label # Node label
nodes1$size <- nodes1$audience.size # Node size
nodes1$borderWidth <- 2 # Node border width

nodes1$color.background <- c("slategrey", "tomato", "gold")[nodes1$media.type]
nodes1$color.border <- "black"
nodes1$color.highlight.background <- "orange"
nodes1$color.highlight.border <- "darkred"

visNetwork(nodes1, links1)

# Change around some properties
links1$width <- 1+links1$weight/8 # line width
links1$color <- "gray"    # line color  
links1$arrows <- "middle" # arrows: 'from', 'to', or 'middle'
links1$smooth <- FALSE    # should the edges be curved?
links1$shadow <- FALSE    # edge shadow

visNetwork(nodes1, links1)

#add a dropdown! 
visNetwork(nodes1, links1) %>%
  visOptions(highlightNearest = TRUE, 
             selectedBy = "type.label")
```

